---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, include=FALSE}


# =======================================================================================
# Combine_data.Rmd
# 
# 14/06/2021 First coding 
# =======================================================================================

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

options(dplyr.summarise.inform = FALSE)

library(tidyverse)
library(lubridate)
require(RJSONIO)  # for finding dropbox folder

source("theme_publication.r")
source("lowcase.r")
source("get_dropbox.r")
source("loadRDataObject.r")
source("encode_zchords.r")

dropboxdir <- file.path(get_dropbox(), "SPFA-PFA_PandoraData")

#load generic data
load(file.path(dropboxdir,"rdata","world.df.RData"))
load(file.path(dropboxdir,"rdata","fao.df.RData"))

load(file.path(dropboxdir,"rdata","icesrectangles.df.RData"))
icesrectangles.df <-
  icesrectangles.df %>% 
  group_by(rect) %>% 
  filter(row_number()==1)

# load haul data
h1 <- 
  loadRData(file.path(dropboxdir, "PFA data","pfa_20210413_haul_extraction_for_pandora.RData")) %>% 
  mutate(
    source="pfa",
    year    = year(date),
    month   = month(date),
    week    = week(date),
    hid    = paste("pfa", tid, stringr::str_pad(haul, pad="0", width=3), sep="_")
  ) %>% 
  rename(
    datetime = haul_time
  )

h2 <- 
  loadRData(file.path(dropboxdir, "SPFA data","spfa_ss_haul.RData")) %>% 
  rename(
    shootlon = londd, 
    shootlat = latdd, 
    date     = dte, 
    vid      = vssl, 
    catch    = est_catch, 
    species  = sp
  ) %>% 
  mutate(
    source  = "spfa", 
    species = tolower(species),
    tid     = paste(vid, stringr::str_pad(trip, pad="0", width=3), sep="_"),
    hid     = paste("spfa", un_hl_id),
    month   = month(date),
    week    = week(date)
  )

# janitor::compare_df_cols(h1, h2)
# combine
h <- 
  bind_rows(h1, h2) %>% 
  dplyr::select(
    vid, tid, hid, date, datetime, year, month, week, 
    shootlat, shootlon, rect,
    source, species, catch
  )


# read length data
l1 <- 
  loadRData(file.path(dropboxdir, "PFA data","pfa_20210413_length_extraction_for_pandora.RData")) %>% 
  mutate(
    source="pfa",
    year    = year(date),
    month   = month(date),
    week    = week(date),
    hid    = paste("pfa", tid, stringr::str_pad(haul, pad="0", width=3), sep="_")
  ) %>% 
  rename(
    datetime = haul_time
  ) %>% 
  ungroup()

l2 <- 
  loadRData(file.path(dropboxdir, "SPFA data","spfa_ss_ltfreq.RData")) %>% 
  rename(
    length = Length_cm, 
    date     = dte, 
    vid      = vssl, 
    species  = sp
  ) %>% 
  mutate(
    source  = "spfa", 
    species = tolower(species),
    tid     = paste(vid, stringr::str_pad(trip, pad="0", width=3), sep="_"),
    hid     = paste("spfa", un_hl_id),
    month   = month(date),
    week    = week(date)
  ) %>% 
  left_join(dplyr::select(h2, hid, shootlat, shootlon)) %>% 
  ungroup()

# janitor::compare_df_cols(l1, l2)
l <- 
  bind_rows(l1, l2) %>% 
  dplyr::select(
    vid, tid, hid, date, datetime, year, month, week, 
    shootlat, shootlon, rect,
    source, species,
    length, count
  )


# bio data - TO BE DONE
b1 <- 
  loadRData(file.path(dropboxdir, "PFA data","pfa_20210413_bio_extraction_for_pandora.RData"))
b2 <- 
  loadRData(file.path(dropboxdir, "SPFA data","spfa_ss_bio.RData"))


  



```

PANDORA working document xx

**Comparison of SPFA and PFA self-sampling data**

Martin Pastoors, Katie Bridgen, Steve Mackinson

`r format(Sys.time(), '%d/%m/%Y')`

**Introduction**

**Data and methods**

**Results**

Comparison of catches of herring, mackerel and blue whiting by year and per haul, for SPFA and PFA separately. 

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

invisible(gc())
xrange <- range(h$shootlon, na.rm=TRUE)
yrange <- c(min(h$shootlat, na.rm=TRUE),64)
yrange <- c(50,62)

tt <-
  h %>% 
  filter(species %in% c("her","mac","whb")) %>% 
  filter(year %in% 2018:2020) %>% 
  filter(shootlon >= xrange[1] & shootlon <= xrange[2]) %>% 
  filter(shootlat >= yrange[1] & shootlat <= yrange[2]) %>% 
  group_by(year, species, source) %>% 
  summarise(n = n())

h %>% 
  filter(species %in% c("her","mac","whb")) %>% 
  filter(year %in% 2018:2020) %>% 
  filter(shootlon >= xrange[1] & shootlon <= xrange[2]) %>% 
  filter(shootlat >= yrange[1] & shootlat <= yrange[2]) %>% 
  
  ggplot(aes(shootlon, shootlat)) + 
  theme_publication() +
  theme(panel.border     = element_rect(colour="black" , size=0.2),
        panel.grid.major = element_blank(),
        strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm"),
        plot.title       = element_text(vjust=0, size=10, hjust=0),
        axis.text        = element_text(size=6),
        axis.title       = element_blank(),
        legend.key.width = unit(0.5, "cm")) +
  
  # coord_quickmap() +
  coord_quickmap(xlim=xrange , ylim=yrange) +
  geom_polygon(data=world.df, aes(long, lat, group=group), fill = "gray90") +
  geom_point(aes(size=catch, colour=source), alpha = 0.4, shape=20) +
  geom_text(data=filter(tt, source=="pfa"), 
            aes(label = paste("n=", n), colour=source),
            x=-Inf, y=Inf, hjust=0, vjust=1, size=3) +
  geom_text(data=filter(tt, source=="spfa"), 
            aes(label = paste("n=", n), colour=source),
            x=Inf, y=Inf, hjust=1, vjust=1, size=3) +
  scale_size(range = c(0.01,5)) +
  guides(colour = guide_legend(nrow = 1)) + 
  labs(x = NULL, y = NULL, size = "t/haul", title="Catch by species") +
  # facet_wrap(~ month, ncol=6)
  facet_grid(year ~ toupper(species))



```

From the SPFA and PFA self-sampling data, the combinations of year, week and rectangles were both SPFA and PFA carried out length-frequency measurements were selected and the aggregated length compositions are shown in the facets below. 

\newpage

Comparison of Herring length compositions by shared week_rectangle combinations

```{r, echo=FALSE, fig.asp=0.8125, fig.align="center", message=FALSE, warning=FALSE}

# myspecies <- "her"

plot_week_rectangle <- function (myspecies ) {
  
  # rectangle/week combinations with both PFA and SPFA length observations
  t <-
    l %>% 
    mutate(year_rect_week = paste(year, rect, week, sep="_")) %>% 
    filter(species == myspecies) %>% 
    group_by(source, species, year_rect_week) %>%
    summarise(count = sum(count, na.rm=TRUE)) %>% 
    group_by(species, year_rect_week) %>%
    mutate(nsources = n()) %>% 
    filter(nsources == 2) 
  
  tt <-
    t %>% 
    group_by(source, species, year_rect_week) %>%
    summarise(n = as.integer(sum(count, na.rm=TRUE)))
  
  # Example of LF by lat/long and year
  l %>% 
    
    # filter(!is.na(shootlon), !is.na(shootlat)) %>% 
    mutate(year_rect_week = paste(year, rect, week, sep="_")) %>% 
    filter(year_rect_week %in% t$year_rect_week) %>% 
    
    left_join(dplyr::select(icesrectangles.df, rect, lat, long), 
              by="rect") %>% 
    filter(species == myspecies) %>% 
    filter(year >= 2018) %>% 
    
    # filter(year_rect_week == "2018_48E9_30") %>% 
    # View()
  # mutate(rect   = encode_zchords(shootlon, shootlat,dx = 0.5,dy = 0.5)) %>%
    # separate(rect, c("shootlon", "shootlat"), sep = ":", convert = TRUE, remove = FALSE) %>% 
    
    mutate(length = floor(length)) %>% 
    
    # summarize by source, year_rect_week
    group_by(year_rect_week, rect, lat, long, source, species, length) %>% 
    summarise(count=sum(count, na.rm=TRUE)) %>% 
    group_by(year_rect_week, rect, lat, long, source, species) %>% 
    mutate(prop=count/sum(count, na.rm=TRUE)) %>% 
    
    ggplot(aes(length, prop)) + 
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          panel.grid.major = element_blank(),
          panel.spacing    = unit(0.5,"mm"),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(vjust=0, size=10),
          axis.text        = element_text(size=6),
          axis.title       = element_blank(), 
          legend.key.width = unit(0.5, "cm")) +
    geom_line(aes(colour=source), alpha = 0.4) +
    geom_text(data=filter(tt, source=="pfa"), 
              aes(label = paste("n=", n), colour=source),
              x=-Inf, y=Inf, hjust=0, vjust=1, size=3) +
    geom_text(data=filter(tt, source=="spfa"), 
              aes(label = paste("n=", n), colour=source),
              x=Inf, y=Inf, hjust=1, vjust=1, size=3) +
    guides(colour = guide_legend(nrow = 1)) + 
    labs(x = NULL, y = NULL, size = "t/haul", title=toupper(myspecies)) +
    # facet_wrap(~ month, ncol=6)
    facet_wrap(~ year_rect_week, ncol=8)
  
}


plot_week_rectangle(myspecies="her" ) 



```

\newpage

Comparison of Mackerel length compositions by shared week_rectangle combinations

```{r, echo=FALSE, fig.asp=1.3, fig.align="center", message=FALSE, warning=FALSE}

plot_week_rectangle(myspecies="mac" ) 

```

\newpage

Comparison of Blue whiting length compositions by shared week_rectangle combinations

```{r, echo=FALSE, fig.asp=0.4875, fig.align="center", message=FALSE, warning=FALSE}

plot_week_rectangle(myspecies="whb" ) 

```

\newpage

When looking for details on some of the year-rectangle-week combinations that showed different length compositions between SPFA and PFA, we are here zooming in on Herring length compositions per haul by selected shared week_rectangle combinations. 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# myspecies <- "her"
# myspecies <- "her"; myyear_rect_week <- "2018_48E9_30"

plot_week_rectangle_byhaul <- function (myspecies, myyear_rect_week ) {
  
  # myspecies <- "her"; myyear_rect_week <- "2018_48E9_30"
  # myspecies <- "whb"; myyear_rect_week <- "2020_35D5_11"
  
  t <-
    l %>% 
    mutate(year_rect_week = paste(year, rect, week, sep="_")) %>% 
    filter(species == myspecies) %>% 
    filter(year_rect_week %in% myyear_rect_week) %>%
    
    mutate(length = floor(length)) %>% 
    
    group_by(source, species, year_rect_week, vid, hid, length) %>%
    summarise(count = sum(count, na.rm=TRUE)) %>% 
    
    group_by(source, species, year_rect_week, vid, hid) %>%
    mutate(prop = count/sum(count, na.rm=TRUE)) 
  
  tt <-
    t %>% 
    group_by(source, species, year_rect_week) %>% 
    summarise(
      nhaul = n_distinct(hid),
      nvessel = n_distinct(vid)
    )
  
  t %>% 
    ggplot(aes(length, prop, group=hid)) + 
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          panel.grid.major = element_blank(),
          panel.spacing    = unit(0.5,"mm"),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(vjust=0, size=10),
          axis.text        = element_text(size=6),
          axis.title       = element_blank(), 
          legend.key.width = unit(0.5, "cm"),
          legend.position  = "none") +
    geom_line(aes(colour=vid), alpha = 0.4) +
    geom_text(data=filter(tt, source=="pfa"), 
              aes(label = paste("nhauls=", nhaul,"\nnvessels=",nvessel)),
              x=-Inf, y=Inf, hjust=0, vjust=1, size=3, inherit.aes = FALSE) +
    geom_text(data=filter(tt, source=="spfa"), 
              aes(label = paste("nhauls=", nhaul,"\nnvessels=",nvessel)),
              x=Inf, y=Inf, hjust=1, vjust=1, size=3, inherit.aes = FALSE) +
    guides(colour = guide_legend(nrow = 1)) + 
    labs(x = NULL, y = NULL, title=paste0(toupper(myspecies), 
                                          " comparison by rect_week_haul")) +
    # facet_wrap(~ month, ncol=6)
    facet_grid(year_rect_week ~ source)
  
}


plot_week_rectangle_byhaul(myspecies="her", 
                           myyear_rect_week <- c("2018_48E9_30","2020_47E8_34", "2020_49E9_30",
                                                 "2020_48E9_24")) 



```

The year-rectangle-week combination 2020_48E9_24 displays an interesting combination of different length compositions. In this case the PFA length compositions are generate by 3 different vessels. The blue length distributions appear to be quite distinct from the other length compositions. Could this be due to different fishing locations? 

\newpage

The year-rectangle-week combinations that were shown in the previous figure can also be assessed for the specific locations where the hauls took place that generated the length compositions. The plot is showing the Herring haul positions by selected shared week_rectangle combinations. 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# myspecies <- "her"
# myspecies <- "her"; myyear_rect_week <- "2018_48E9_30"


plot_week_rectangle_byhaul_bylatlong <- function (myspecies, myyear_rect_week ) {
  
  # myspecies <- "her"; myyear_rect_week <- "2018_48E9_30"
  # myspecies <- "whb"; myyear_rect_week <- "2020_35D5_11"
  
  t <-
    l %>% 
    mutate(year_rect_week = paste(year, rect, week, sep="_")) %>% 
    filter(species == myspecies) %>% 
    filter(year_rect_week %in% myyear_rect_week) %>%
    
    mutate(length = floor(length)) %>% 
    
    group_by(source, species, year_rect_week, vid, hid, shootlat, shootlon) %>%
    summarise(count = sum(count, na.rm=TRUE))  
  
  tt <-
    t %>% 
    group_by(source, species, year_rect_week) %>% 
    summarise(
      nhaul = n_distinct(hid),
      nvessel = n_distinct(vid)
    )
  
  xrange <- range(t$shootlon, na.rm=TRUE)
  yrange <- range(t$shootlat, na.rm=TRUE)
  
  t %>% 
    ggplot(aes(shootlon, shootlat, group=hid)) + 
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          # panel.grid.major = element_blank(),
          panel.grid.major = element_line(colour="black" , size=0.2),
          panel.spacing    = unit(0.5,"mm"),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(vjust=0, size=10),
          axis.text        = element_text(size=6),
          axis.title       = element_blank(), 
          legend.key.width = unit(0.5, "cm"),
          legend.position  = "none") +
    coord_quickmap(xlim=xrange , ylim=yrange) +
    geom_point(aes(colour=vid, size=count), alpha = 0.4) +
    geom_text(data=filter(tt, source=="pfa"), 
              aes(label = paste("nhauls=", nhaul,"\nnvessels=",nvessel)),
              x=-Inf, y=Inf, hjust=0, vjust=1, size=3, inherit.aes = FALSE) +
    geom_text(data=filter(tt, source=="spfa"), 
              aes(label = paste("nhauls=", nhaul,"\nnvessels=",nvessel)),
              x=Inf, y=Inf, hjust=1, vjust=1, size=3, inherit.aes = FALSE) +
    guides(colour = guide_legend(nrow = 1)) + 
    labs(x = NULL, y = NULL, title=paste0(toupper(myspecies), 
                                          " haul positions compared by rect_week_haul")) +
    scale_x_continuous(breaks = seq(-10,10,by = 1)) +
    scale_y_continuous(breaks = seq(50,70,by = 0.5)) +
    # facet_wrap(~ month, ncol=6)
    facet_grid(year_rect_week ~ source)
  
}


plot_week_rectangle_byhaul_bylatlong(
  myspecies="her", 
  myyear_rect_week <- c("2018_48E9_30","2020_47E8_34", "2020_49E9_30",
                        "2020_48E9_24")) 

```

Again focusing on the 2020_48E9_24 combination, the blue hauls are quite close to the other hauls in that rectangle. This type of result requires a more indepth look at the basis of the length-estimates. 
 